---
title: "An Analytics Workflow of Happiness Data"
author: "Molly Izenson"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
    number_sections: FALSE
    theme: spacelab
    code_folding: hide

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Analysis Overview
This data comes from the 2020 World Happiness Report, a survey that ranks 156 countries on how happy their citizens perceive themselves to be. The most important source of data from the Report is the Gallup World Poll, which surveys citizens' life evaluations, leading to the happiness ranking. 

The report can be found [here](https://worldhappiness.report/ed/2020/).

### Goal of Analysis
The goal of this analysis is to determine which factors have a statistically significant impact on a country's happiness, using the variable, Life Ladder, as y. This analysis is to be used by the World Health Organization (WHO) to better understand why and help countries who may have low levels of happiness. This analysis should provide insight to the WHO so that they can institute new policies that might, in the long run, affect happiness. It is crucial that this analysis is reusable in order for the audience to use the knowledge. 

#### Loading Data and Packages into R
```{r}
happy<-read.csv("HappyData.csv")
library(pacman)
pacman::p_load(dplyr, coefplot, DataExplorer)
```
* The dplyr package allows for manipulation of variables, such as selecting variables to drop from the analysis.
* The DataExplorer package gives the ability to introduce the data and create graphs to easily visualize the data.
* The coefplot is used to visualize the confidence intervals of the coefficients in the regression model.

# Description of Data
There are 1848 observations with 26 variables.A full dictionary of the variables can be found [here](https://happiness-report.s3.amazonaws.com/2020/WHR20_Ch2_Statistical_Appendix.pdf).

* __Life.Ladder__ or happiness score for a country on a given year, on a scale of 0 to 10.
 * __Country.name__ is the name of 156 countries.
 * __Year__ is the year the data was collected, from 2005 to 2019.
 * __Log.GDP.per.capita__ is the log of a country's GDP per capita.
 * __Healthy.life.expectancy.at.birth__ is the average life expectancy for a person at birth.
 * __Social.Support__ is the percentage of people who say they have someone to count on in times of trouble 
 * __Freedom.to.make.life.choices__ is the national average of a citizen's belief on their amount of freedom.
 * __Generosity__ is the residual of regressing national average to the GWP response question, "Have you donated money to a charity in the past month? on GDP per capita.
 * __Perceptions.of.Corruption__ is the percentage of people who believe there is corruption in their country, both within government and businesses.
 * __Positive.Affect__ is the average of how many times a person feels happiness, laugh, and enjoyment in a day.
 * __Negative Affect__ is the average of how many times a person feels worry, sadness, anger in a day.
 * __giniIncGallup__ measures the GINI of household income.
 * __giniIncWB__ is the GINI index from the World Bank.
 * __Average GINI index__ from the World Bank (variable = giniIncWBavg)
 * __Standard.deviation.of.ladder.by.country.year__ measures the inequality of happiness.
 * __Standard.deviation.mean.of.ladder.year.by.country.year__ measures the mean of the inequality of happiness.
 * __Democratic.Quality__ measures how democratic a country's government is. 
 * __Delivery.Quality__ measures how democratic a country's government is.
 * __GINI.index..World.Bank.estimate__ measures the World Bank data for a country.
 * __Most.people.can.be.trusted..WVS.round.1981.1984__ is the average percentage of people in a country that most people can be trused from 1981 to 1984.
 * __Most.people.can.be.trusted..WVS.round.1989.1993__ is the average percentage of people in a country that most people can be trused from 19889 to 1993.
 * __Most.people.can.be.trusted..WVS.round.1994.1998__ is the average percentage of people in a country that most people can be trused from 1994 to 1998.
 * __Most.people.can.be.trusted..WVS.round.1999.2004__ is the average percentage of people in a country that most people can be trused from 1999 to 2004.
 * __Most.people.can.be.trusted..WVS.round.2005.2009__ is the average percentage of people in a country that most people can be trused from 2005 to 2009.
 * __Most.people.can.be.trusted..WVS.round.2010.2014__ is the average percentage of people in a country that most people can be trused from 2010 to 2014.
 * __Most.people.can.be.trusted..Gallup__ is the average percentage of people in a country that most people can be trused.
 
## Introduction of Data 
The head and tails of the data below give the first and last six observations for the first five variables in the dataset. As you can see, there are multiple years recorded for each country.
```{r}
head(happy)[1:5]
tail(happy)[1:5]
``` 

```{r}
introduce(happy)
```
As mentioned above, there are 1848 observations from 26 variables, with 12,297 missing values. The missing values will be addressed further along in this analysis.

```{r}
plot_histogram(happy)
```

These histograms give insight quick visual insight into the remaining variables. It is clear that some of the variables are skewed, such as "Perceptions.of.corruption" and "Social.support", meaning the distribution of these variables varies from a normal distribution. 

## Missing Data
```{r}
plot_missing(happy)
```

As you can see from the chart above and the data summary, there are 12,297 missing values in this data set.The plot above shows the percentages of missing data for each variable, determining if it is okay to keep it in, or if it should be removed.

The Happiness Report acknowledges these missing values and says they did not use missing value to draw results from the happiness rankings and the regression outputs. However, they did make use of imputation in decomposing a country's average ladder score into components explained by 6 determinants (GDP per person, healthy life expectancy, social support, perceived freedom, generosity, and perception of corruption).

A decision must be made on whether to keep or remove these variables with high percentages of missing values in the analysis. I am going to remove all variables with "Bad" or "Remove" labels.

# Cleaning the Data
The code below removes the 8 variables with missing data considered "Bad" or "Remove" from the dataset. Removing these variables will allow for the analysis to be more accurate.
```{r}
library(dplyr)
happy<-select(happy, -Most.people.can.be.trusted..WVS.round.2010.2014, -GINI.index..World.Bank.estimate., -Most.people.can.be.trusted..WVS.round.2005.2009, -Most.people.can.be.trusted..WVS.round.1994.1998, -Most.people.can.be.trusted..WVS.round.1999.2004, -Most.people.can.be.trusted..WVS.round.1989.1993, -Most.people.can.be.trusted..Gallup, -Most.people.can.be.trusted..WVS.round.1981.1984)

```

```{r}
introduce(happy)
```
Now there are 1385 missing values. 

## Imputation of Variables
In order to account for the 1385 missing values, instead of choosing to eliminate each variables with missing and losing the potentially important information contained, I will impute each variable with the median for each variable. By choosing median over mean, it  will limit the influence of outliers. After the imputation, I will remove the indicator variable in order to reduce my degrees of freedom in the model.

```{r, echo=FALSE, results='hide'}
#gini.of.household.income.reported.in.Gallup..by.wp5.year
summary(happy$gini.of.household.income.reported.in.Gallup..by.wp5.year)
happy$M_gini.of.household.income.reported.in.Gallup..by.wp5.year<-as.factor(ifelse(is.na(happy$gini.of.household.income.reported.in.Gallup..by.wp5.year), 1, 0))
happy$gini.of.household.income.reported.in.Gallup..by.wp5.year[is.na(happy$gini.of.household.income.reported.in.Gallup..by.wp5.year)]<-median(happy$gini.of.household.income.reported.in.Gallup..by.wp5.year, na.rm=TRUE)
summary(happy$gini.of.household.income.reported.in.Gallup..by.wp5.year)
```

```{r, echo=FALSE, results='hide'}
#Imputation for confidence in national government:
summary(happy$Confidence.in.national.government)
happy$M_confidence.in.national.government<-as.factor(ifelse(is.na(happy$Confidence.in.national.government), 1, 0))
happy$Confidence.in.national.government[is.na(happy$Confidence.in.national.government)]<-median(happy$Confidence.in.national.government, na.rm=TRUE)
summary(happy$Confidence.in.national.government)
```

```{r, echo=FALSE, results='hide'}
#imputation for GINI.index..World.Bank.estimate...average.2000.2017..unbalanced.panel
summary(happy$GINI.index..World.Bank.estimate...average.2000.2017..unbalanced.panel)
happy$M_GINI.index..World.Bank.estimate...average.2000.2017..unbalanced.panel<-as.factor(ifelse(is.na(happy$GINI.index..World.Bank.estimate...average.2000.2017..unbalanced.panel), 1, 0))
happy$GINI.index..World.Bank.estimate...average.2000.2017..unbalanced.panel[is.na(happy$GINI.index..World.Bank.estimate...average.2000.2017..unbalanced.panel)]<-median(happy$GINI.index..World.Bank.estimate...average.2000.2017..unbalanced.panel, na.rm=TRUE)
summary(happy$GINI.index..World.Bank.estimate...average.2000.2017..unbalanced.panel)
```

Here is an example of the process for imputation for the variable "Democratic.Quality". The same process was repeated across the 13 variables that had any missing observations.
```{r}
#imputation for Democratic.Quality
summary(happy$Democratic.Quality)
happy$M_Democratic.Quality<-as.factor(ifelse(is.na(happy$Democratic.Quality), 1, 0))
happy$Democratic.Quality[is.na(happy$Democratic.Quality)]<-median(happy$Democratic.Quality, na.rm=TRUE)
summary(happy$Democratic.Quality)
```

```{r, echo=FALSE, results='hide'}
#imputation for DeliveryQuality
summary(happy$Delivery.Quality)
happy$M_Delivery.Quality<-as.factor(ifelse(is.na(happy$Delivery.Quality), 1, 0))
happy$Delivery.Quality[is.na(happy$Delivery.Quality)]<-median(happy$Delivery.Quality, na.rm=TRUE)
summary(happy$Delivery.Quality)
```

```{r, echo=FALSE, results='hide'}
#imputation for perceptions of corruption:
summary(happy$Perceptions.of.corruption)
happy$M_Perceptions.of.corruption<-as.factor(ifelse(is.na(happy$Perceptions.of.corruption), 1, 0))
happy$Perceptions.of.corruption[is.na(happy$Perceptions.of.corruption)]<-median(happy$Perceptions.of.corruption, na.rm=TRUE)
summary(happy$Perceptions.of.corruption)
```

```{r, echo=FALSE, results='hide'}
#imputation for generosity
summary(happy$Generosity)
happy$M_Generosity<-as.factor(ifelse(is.na(happy$Generosity), 1, 0))
happy$Generosity[is.na(happy$Generosity)]<-median(happy$Generosity, na.rm=TRUE)
summary(happy$Generosity)
```

```{r, echo=FALSE, results='hide'}
#imputation for healthy life expectancy at birth:
summary(happy$Healthy.life.expectancy.at.birth)
happy$M_Healthy.life.expectancy.at.birth<-as.factor(ifelse(is.na(happy$Healthy.life.expectancy.at.birth), 1, 0))
happy$Healthy.life.expectancy.at.birth[is.na(happy$Healthy.life.expectancy.at.birth)]<-median(happy$Healthy.life.expectancy.at.birth, na.rm=TRUE)
summary(happy$Healthy.life.expectancy.at.birth)
```

```{r, echo=FALSE, results='hide'}
#imputation for freedom to make life choices
summary(happy$Freedom.to.make.life.choices)
happy$M_Freedom.to.make.life.choices<-as.factor(ifelse(is.na(happy$Freedom.to.make.life.choices), 1, 0))
happy$Freedom.to.make.life.choices[is.na(happy$Freedom.to.make.life.choices)]<-median(happy$Freedom.to.make.life.choices, na.rm=TRUE)
summary(happy$Freedom.to.make.life.choices)
```

```{r, echo=FALSE, results='hide'}
#imputation for Log.GDP per capita
summary(happy$Log.GDP.per.capita)
happy$M_Log.GDP.per.capita<-as.factor(ifelse(is.na(happy$Log.GDP.per.capita), 1, 0))
happy$Log.GDP.per.capita[is.na(happy$Log.GDP.per.capita)]<-median(happy$Log.GDP.per.capita, na.rm=TRUE)
summary(happy$Log.GDP.per.capita)
```

```{r, echo=FALSE, results='hide'}
#imputation for positive affect
summary(happy$Positive.affect)
happy$M_Positive.affect<-as.factor(ifelse(is.na(happy$Positive.affect), 1, 0))
happy$Positive.affect[is.na(happy$Positive.affect)]<-median(happy$Positive.affect, na.rm=TRUE)
summary(happy$Positive.affect)
```

```{r, echo=FALSE, results='hide'}
#imputation for negative affect:
summary(happy$Negative.affect)
happy$M_Negative.affect<-as.factor(ifelse(is.na(happy$Negative.affect), 1, 0))
happy$Negative.affect[is.na(happy$Negative.affect)]<-median(happy$Negative.affect, na.rm=TRUE)
summary(happy$Negative.affect)
```

```{r, echo=FALSE, results='hide'}
#imputation for social support
summary(happy$Social.support)
happy$M_Social.support<-as.factor(ifelse(is.na(happy$Social.support), 1, 0))
happy$Social.support[is.na(happy$Social.support)]<-median(happy$Social.support, na.rm=TRUE)
summary(happy$Social.support)
```

Now, we can remove the variables created for the imputation containing the median. 
```{r, echo=FALSE}
#remove M columns after imputation
happy <- select(happy, -M_Social.support, -M_Negative.affect, -M_Positive.affect, -M_Log.GDP.per.capita, -M_Freedom.to.make.life.choices, -M_Healthy.life.expectancy.at.birth, -M_Generosity, -M_Perceptions.of.corruption, -M_Delivery.Quality, -M_Democratic.Quality, -M_GINI.index..World.Bank.estimate...average.2000.2017..unbalanced.panel, -M_confidence.in.national.government, -M_gini.of.household.income.reported.in.Gallup..by.wp5.year) 

```

As you can see, now there are zero missing observations in the dataset.
```{r}
plot_missing(happy)
```

With no missing data, we can look at the correlation between variables. 
```{r, echo=FALSE}
happy$year <- as.integer(happy$year)
```

```{r, echo=FALSE}
happy_numeric<-unlist(lapply(happy, is.numeric))
happy_numeric2<-happy[, happy_numeric]
```

```{r}
M<-cor(happy_numeric2, use = "complete.obs")
library("corrplot")
corrplot(M, type = "upper", tl.pos = "td",
         method = "circle", tl.cex = 0.5, tl.col = 'black',
         order = "hclust", diag = TRUE)
```

The correlation plot is an easy way to visualize associations between variables. Strong positive associations are indicated by large dark blue dots and strong negative associations are indicated by large dark red dots.

From the correlation plot, we can determine that there is a strong positive associations between the "standard.deviation.of.ladder.by.country.year" and the "standard.deviation.mean.of.ladder.by.country.year" and a strong negative association between "standard.deviation.mean.of.ladder.by.country.year" and "Life.ladder." So, I will not include  will not include either "standard.deviation.mean.of.ladder.by.country.year" or "standard.deviation.of.ladder.by.country.year" in the regression analysis because of its strong association with the target variable, which could lead to skewed results. It makes sense that these two variables would be correlated because the nature of the variable is directly related to the Ladder variable.

The correlation plot also shows a strong association between Delivery.Quality and Democratic.Quality. It only makes sense to include one of these variables since they contain the same information. I chose to include Democratic.Quality.

```{r}
happy<-select(happy, -Standard.deviation.of.ladder.by.country.year, -Standard.deviation.Mean.of.ladder.by.country.year, -Delivery.Quality)
```

# The Regression Model
Now that the data is clean and preprocessed, it is time to consider the type of model to use to determine what factors effect a country's happiness. I will use an explanatory regression model because we are interested in identifying the factors that affect a country's happiness score. I will use a multiple linear regression model with interaction terms for the analysis.

There are several varaibles not included in the analysis because it would not add to the model, including Country.Name and year. The results of the model are below. 
```{r}
happy.reg<-lm(Life.Ladder~Log.GDP.per.capita+Social.support+Healthy.life.expectancy.at.birth+Freedom.to.make.life.choices+Generosity+Perceptions.of.corruption+Positive.affect+Confidence.in.national.government+gini.of.household.income.reported.in.Gallup..by.wp5.year+-Country.name, data=happy)
summary(happy.reg)
```
The explanatory regression model includes 12 variables. All of the variables included in the model are significant with p-values above alpha = .05.

We would expect all of the variables that are included in the model to make sense. For example, it is reasonable to assume that the more freedom a citizen has to make life choices, their happiness would increase, therefore increasing their country's happiness score. Similarly, the would expect as the percentage of people in a country that believe that their government has widespred corruption decreases, the happiness score also decreases.

```{r, echo=FALSE, results='hide'}
confint(happy.reg,level=0.95)
```

# Visualizations of Model
```{r}
coefplot(happy.reg, xlab = "Coefficient Confidence Intervals", varnames=NULL, intercept=FALSE)
```

__Coefficient Plot Analysis:__
The coefficient plot above displays the point estimates for each variable in the model and their confidence intervals.For example, if we wanted to interpret the coefficient of "Social.Support" we could say the mean of the happiness score (Life.Ladder variable) increases by 1.665423 points given a one-unit increase in the Social.Support score while holding other variables in the model constant.

# Justification of Model

__Slope interpretation:__
If all the variables in the model are scored as 0, the Life.Ladder, or happiness, score is estimated to be 4.159059 (on a 0-10 scale).

__Model significance test:__

  $H0: β1=β2=β3=β4=β5=β6=β7=β8=β9=β10=β11=β12=β13=0$
  
  Ha: at least one $βk≠0$

The F-statistic is 561.7, with degrees of freedom k=12 and n−(k+1)=1835.
The p-value is < 2.2e-16. 

There is significant evidence at any reasonable level of significance to conclude that any of the variables in the model are significant linear predictors of the happiness level of a country.

__$R^2$  Interpretation:__
An $R^2$ score of 0.786 (Adjusted R-squared: 0.7846 accounts for the sample size and model parameters) implies that 78.6% of the observed variation in the happiness score of countries are accounted for by the model that includes a combination of the 13 variables in the model.

## Plot of Assumptions
Residuals vs. Fitted Values:
```{r, echo=FALSE}
plot(happy.reg$fitted.values,happy.reg$residuals,ylab="Residuals",xlab="Fitted Values",main="Residuals vs. Fitted Values")
abline(0,0)
```

The plot of the Residuals vs. Fitted Values above shows the residuals randomly scattered around the 0 line, suggesting that the assumption that the relationship is linear is accurate and that the variances of the error terms are equal. The graph does not show any major outliers.

```{r, echo=FALSE}
qqnorm(happy.reg$residuals)
```

This Normal Q-Q Plot of the residuals appears there is normal distribution in the model, with some slight strays for normalcy in tails of the distribution.

# What does this mean for the World Health Organization?

The explanatory model created in the report is significant, as well as all of the predictor variables included. By determining the most significant factors that influence the Life.Ladder score, the World Health Organization can determine the range of happiness scores by country and explore options for improvement, especially within the variables that have the highest significant influence on the Life.Ladder score. We can use the information on the significant predictors for a country's Life Ladder score in order to identify struggling countrys and create potential solutions.

For example, using the results from the model, the WHO can determine how country's may improve their citizens' Social Support, which if the score improved by 1, could lead to a 1.680946 increase in a country's happiness score, which is a combination of their citizen's perceived happiness. The WHO can explore the variables in the model for each country to gain insight to the country's so that they can institute new policies that might, in the long run, affect happiness.